Hydrodynamics and rheology of adhesive vesicle doublets

### 2019-02-12 vesicle_code/docs/adhesion/adhesion.tex
edited shear example

In Figure~\ref{fig:extensionalInclinationAngle}, we plot the inclination
angle of the doublet as a function of time for four different
extensional rates (left) and the final inclination angle for all
doublets that reach an equilibrium state (middle).  We observe that not
only do smaller extensional rates result in smaller inclination
angles, but smaller reduced areas also result in smaller inclination
angles.  Moreover, the critical Hamaker constant that determines if the
doublet remains trapped or is broken decreases with an increasing
extensional rate \todo[inline]{not sure how we can say this since we've
only considered one Hamaker constant}.  We summarize the final
inclination angle of the doublet in the right plot of
Figure~\ref{fig:extensionalInclinationAngle}.  The size of the round
dots are scaled to the final inclination angle as defined in
Figure~\ref{fig:InclinationAngle}.  At smaller extensional rates, the
vesicles come closer to aligning their long axis with the diverging
direction.  When the doublet is broken at reduced area $\Delta A =
0.80$, the extensional rate is sufficiently large to align the long axis
of the vesicle with the diverging direction and then the vesicles
separate (Figure~\ref{fig:extensional3}).  For very low extensional
rates, the time horizon is insufficient for the doublet to reach an
equilibrium state, and these simulations are marked with a square.  The
right plot of Figure~\ref{fig:extensionalInclinationAngle} also
summarizes the reduced areas and extensional rates that result in a
bound vesicle doublet in a fluid trap: Parameter values with a blue mark
result in a fluid trap and parameter values with a red mark result in
vesicle separation.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Adhesion of two vesicles in a shear flow}
\label{sec:sflow} 
We consider two vesicles suspended in the shear flow $\uu = \chi(y,0)$,
where $\chi$ is the shear rate.  The vesicles are placed on the $y$ axis
so that the background velocity drives the vesicles towards one another.
In the absence of adhesion, once the vesicles are sufficiently close,
they are deflected to opposite sides of the $y$ axis, pass one another,
and then separate.  However, in the presence of adhesion, the vesicles
can form a doublet for certain values of the separation distance
$\delta$, Hamaker constant $\mathcal{H}$, shear rate $\chi$, and reduced
area $\Delta A$.  Figure~\ref{fig:doublet090} shows snapshots of two
vesicles that have formed a doublet and the color coding is the tension
along the vesicle.  The two vesicles are identical with reduced area
$\Delta A = 0.9$, shear rate $\chi = 0.5$, separation distance $\delta =
0.4$, and Hamaker constant $\mathcal{H} = 0.7$.  The vesicles in the
doublet move in tandem rather than performing the tank-treading motion
that is observed in the absence of adhesion.  Similar to the
quiescent example, the membrane tension is negative in the contact
region indicating that the membrane is being compressed when the
adhesive force is strongest.

\begin{figure}[htp]
  \includegraphics[width=0.24\textwidth]{figs/adR4em1adS7em1Chi5em1_ra090_image01.png}
  \includegraphics[width=0.24\textwidth]{figs/adR4em1adS7em1Chi5em1_ra090_image02.png}
  \includegraphics[width=0.24\textwidth]{figs/adR4em1adS7em1Chi5em1_ra090_image03.png}
  \includegraphics[width=0.24\textwidth]{figs/adR4em1adS7em1Chi5em1_ra090_image04.png}
  \includegraphics[width=0.24\textwidth]{figs/adR4em1adS7em1Chi5em1_ra090_image05.png}
  \includegraphics[width=0.24\textwidth]{figs/adR4em1adS7em1Chi5em1_ra090_image06.png}
  \includegraphics[width=0.24\textwidth]{figs/adR4em1adS7em1Chi5em1_ra090_image07.png}
  \includegraphics[width=0.24\textwidth]{figs/adR4em1adS7em1Chi5em1_ra090_image08.png}
  \caption{\label{fig:doublet090} The formation of a doublet in a shear
  flow.  The reduced area is $\Delta A = 0.9$, the Hamaker constant is
  $\mathcal{H}=0.7$, the adhesion length scale is $\delta = 0.4$, and
  the shear rate is $\chi=0.5$.}
\end{figure}

We characterize the effect of the Hamaker constant and shear rate by
determining a critical Hamaker constant, $\mathcal{H}_c$, which
determines if the vesicles form a doublet or separate.  We fix the
adhesion length scale to $\delta = 0.4$ and the vesicle reduced area to
$\Delta A = 0.9$.  In the left plot of
Figure~\ref{fig:sflow_phase_diagram}, we plot the minimum distance
between the vesicles as a function of time for various Hamaker
constants.  The red curves correspond to Hamaker constants that are too
small to bind the vesicles into a doublet, and the blue curves
correspond to Hamaker constants that are sufficiently large to form a
doublet.

\begin{figure}
  \includegraphics[height=0.35\textwidth]{figs/shear_adR4em1Chi1e0_ra-090.pdf}
  \includegraphics[height=0.35\textwidth]{figs/shear_adR4em1_ra090_phaseDiagram.pdf}
  \caption{{\em Left}: Distance between a pair of vesicles in a planar
  shear flow with shear rate $\chi=0.5$ and adhesion length scale
  $\delta = 0.4$.   The different lines correspond to a linear spacing
  of adhesion strengths ranging from $\mathcal{H}=0.1$ and
  $\mathcal{H}=1.0$.  {\em Right}: Phase diagram of vesicle
  hydrodynamics in a shear flow.  At a fixed reduced area $\Delta
  A=0.90$ and $\delta = 0.4$,  the critical Hamaker constant
  $\mathcal{H}$ for binding of two vesicles in a shear flow depends on
  the shear rate $\chi$.
\label{fig:sflow_phase_diagram}}
\end{figure}

Finally, we investigate the rheology of a suspension of a doublet by
computing the effective viscosity of a doublet and compare it to the
effective viscosity of a single tank-treading vesicle. The effective
viscosity is defined as the viscosity of a homogeneous Newtonian fluid
with the same energy dissipation per macroscopic element of fluid.  In a
simple shear flow, the intrinsic viscosity, $[\mu]$ is
\begin{align*}
  [\mu]:= \frac{\mu_{\mathrm{eff}} - \mu_0}{\phi \mu_0} = 
  \frac{1}{\chi \mu_0 T} \int_{T_i}^{T_e} 
  \langle \sigma_{12} \rangle dt,
\end{align*}
where
\begin{align*}
  \langle \tau \rangle = \frac{1}{|\omega|} \int_{\gamma}
    \xxi \otimes \xx ds,
\end{align*}
where $\phi$ is the area fraction of vesicles, $\tau$ is the stress due
to the vesicles, and $\langle \cdot \rangle$ is the spatial average.

In the left plot of Figure~\ref{fig:shearIntrinsicViscosity}, we compare
the intrinsic viscosity of a single tank-treading vesicle to a doublet.
To validate our simulations, we superimpose (black marks) the intrinsic
viscosity calculated by Ghigliotti {\em et
al.}~\cite{GhigliottiBibenMisbah2010_JFM}.  The presence of the doublet
significantly increases the intrinsic viscosity at all the reduced
areas.  To further characterize the effect of adhesion, in the right
plot of Figure~\ref{fig:shearIntrinsicViscosity} we decompose the
intrinsic viscosity into the contributions from the bending and tension
(blue), and the contribution from the adhesion (red).  We also
superimpose twice the intrinsic viscosity of a single tank-treading
vesicle (dashed curve) to demonstrate that the bending and tension of
the doublet behave similarly, but not identically, to a dilute
suspensions of non-adhering tank-treading vesicles.  We see that the
effect of the adhesion on the intrinsic viscosity is largest for
vesicles with small reduced areas.

\begin{figure}[htp]
  \centering
  \includegraphics[width=0.45\textwidth]{figs/shear2Ves_adR4em1adS7em1Chi5em1.pdf}
  \includegraphics[width=0.45\textwidth]{figs/doublet_decomp.pdf}
  \caption{\label{fig:shearIntrinsicViscosity} {\em Left}: The intrinsic
  viscosity of a tank-treading vesicle (blue) and a doublet (red).  The
  shear rate is $\chi = 0.5$, the Hamaker constant is $\mathcal{H} =
  0.7$, and the adhesion length scale is $\delta = 0.4$, and the length
  of the vesicles are held fixed.  The black marks denote intrinsic
  viscosity values computed by Ghigliotti {\em et
  al.}~\cite{GhigliottiBibenMisbah2010_JFM} (cf.~Figure 5).  {\em
  Right}: The decomposition of the intrinsic viscosity into the
  contributions for the bending and tension (blue) of the vesicles, and
  the contribution from the vesicle adhesion (red).  Also included is
  twice the intrinsic viscosity of a single tank-treading vesicle
  (dashed black line).}
\end{figure}

@@

Flormann {\it et al.} studied the clustering and packing of vesicles in free suspension~\cite{FlormannAouane2017_SciReports}. 
In engineering applications of vesicle emulsions
often smaller vesicles are enclosed in a big vesicle, as shown in 
Figure~\ref{fig:relaxationManyVes}, where  the left panel shows an initial condition of five vesicles
suspended inside of a larger vesicle.
The second panel from left shows the membrane configuration and flow
streamlines at an intermediate time.  There is no imposed flow, so the
dynamics are governed entirely by the adhesion, bending, and tension
forces.  In the absence of confinement, we would expect the vesicles to
align themselves as a rouleaux with the membranes forming a parachute,
male-female, or S-shape
configurations~\cite{FlormannAouane2017_SciReports}.  However, because
of the bounding vesicle and the dense packing, the interior vesicles
align themselves differently.  In addition, several parts of the
bounding membrane closely follow the interior membranes in order to
minimize the adhesion energy.   As a result of force balance the equilibrium shape of the bounding vesicle
is far from a symmetric shape while the interior smaller vesicles are of similar shape to each other.
We also plot the tension (thrid) and
adhesion (fourth) of the final equilibrium configuration.  
We remark that the adhesion force is all attractive on all the vesicle membranes while large positive
membrane tension is found only on the bounding vesicle. 
%Both the
%vesicle shape, tension, and adhesion qualitatively agree with our
%doublet results.  
In particular, the interior vesicles form a flat
contact region with a large negative tension.
We are now conducting simulations of such vesicle configuration under a planar shear flow.

%\begin{figure}[htp]
%  \centering
%  \includegraphics[width=0.24\textwidth]{figs/relaxationManyVes_Snapshot01.pdf}
%  \includegraphics[width=0.24\textwidth]{figs/relaxationManyVes_Snapshot02.pdf}
%  \includegraphics[width=0.24\textwidth]{figs/relaxationManyVes_Snapshot03.pdf}
%  \includegraphics[width=0.24\textwidth]{figs/relaxationManyVes_Snapshot04.pdf} \\
%  \includegraphics[width=0.24\textwidth]{figs/relaxationManyVes_Snapshot05.pdf}
%  \includegraphics[width=0.24\textwidth]{figs/relaxationManyVes_Snapshot06.pdf}
%  \includegraphics[width=0.24\textwidth]{figs/relaxationManyVes_Snapshot07.pdf}
%  \includegraphics[width=0.24\textwidth]{figs/relaxationManyVes_Snapshot08.pdf}
%  \caption{\label{fig:relaxationManyVes} Five vesicles suspended inside
%  a large vesicle.  The background flow is quiescent, the adhesion
%  length scale is $\delta = 0.4$, and the Hamaker constant is
%  $\mathcal{H} = 0.7$.  At the final snapshot, the suspension is near an
%  equilibrium configuration.}
%\end{figure}

>>>>

In Figure~\ref{fig:extensionalInclinationAngle}, we plot the inclination
angle of the doublet as a function of time for four different
extensional rates (left) and the final inclination angle for all
doublets that reach an equilibrium state (middle).  We observe that not
only do smaller extensional rates result in smaller inclination
angles, but smaller reduced areas also result in smaller inclination
angles.  
%
%From physics we expect that stronger adhesion is required to keep 
%the vesicle doublet intact under a fluid trap as we increase the flow rate.
%
%It is less obvious how the critical adhesion strength depends on the reduced area of the vesicles.
%
%Even though we only used one value of Hamaker constant (${\cal H} = 0.7$) we expect that
%
%Moreover, the critical Hamaker constant that determines if the
%doublet remains trapped or is broken decreases with an increasing
%extensional rate \todo[inline]{not sure how we can say this since we've
%only considered one Hamaker constant}.  
%
We summarize the final
inclination angle of the doublet in the right plot of
Figure~\ref{fig:extensionalInclinationAngle}.  
The size of the round
dots are scaled to the final inclination angle as defined in
Figure~\ref{fig:InclinationAngle}.  At smaller extensional rates, the
vesicles come closer to aligning their long axis with the diverging
direction.  When the doublet is broken at reduced area $\Delta A =
0.80$, the extensional rate is sufficiently large to align the long axis
of the vesicle with the diverging direction and then the vesicles
separate (Figure~\ref{fig:extensional3}).  For very low extensional
rates, the time horizon is insufficient for the doublet to reach an
equilibrium state, and these simulations are marked with a square.  The
right plot of Figure~\ref{fig:extensionalInclinationAngle} also
summarizes the reduced areas and extensional rates that result in a
bound vesicle doublet in a fluid trap: Parameter values with a blue mark
result in a fluid trap and parameter values with a red mark result in
vesicle separation.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Adhesion of two vesicles in a shear flow}
\label{sec:sflow} 
We consider two vesicles suspended in the shear flow $\uu = \chi(y,0)$,
where $\chi$ is the shear rate.  The vesicles are placed on the $y$ axis
so that the background velocity drives the vesicles towards one another.
In the absence of adhesion, once the vesicles are sufficiently close,
they are deflected to opposite sides of the $y$ axis, pass one another,
and then separate.  However, in the presence of adhesion, the vesicles
can form a doublet for certain values of the separation distance
$\delta$, Hamaker constant $\mathcal{H}$, shear rate $\chi$, and reduced
area $\Delta A$.  At all reduced areas, the vesicles have a length of
$2\pi$.  Figure~\ref{fig:doublet090} shows snapshots of two vesicles
that have formed a doublet and the color coding is the tension along the
vesicle.  The two vesicles are identical with reduced area $\Delta A =
0.9$, shear rate $\chi = 0.5$, separation distance $\delta = 0.4$, and
Hamaker constant $\mathcal{H} = 0.7$.  The vesicles in the doublet move
in tandem rather than performing the tank-treading motion that is
observed in the absence of adhesion.  Similar to the quiescent example,
the membrane tension is negative in the contact region indicating that
the membrane is being compressed when the adhesive force is strongest.

\begin{figure}[htp]
  \includegraphics[width=0.24\textwidth]{figs/adR4em1adS7em1Chi5em1_ra090_image01.png}
  \includegraphics[width=0.24\textwidth]{figs/adR4em1adS7em1Chi5em1_ra090_image02.png}
  \includegraphics[width=0.24\textwidth]{figs/adR4em1adS7em1Chi5em1_ra090_image03.png}
  \includegraphics[width=0.24\textwidth]{figs/adR4em1adS7em1Chi5em1_ra090_image04.png}
  \includegraphics[width=0.24\textwidth]{figs/adR4em1adS7em1Chi5em1_ra090_image05.png}
  \includegraphics[width=0.24\textwidth]{figs/adR4em1adS7em1Chi5em1_ra090_image06.png}
  \includegraphics[width=0.24\textwidth]{figs/adR4em1adS7em1Chi5em1_ra090_image07.png}
  \includegraphics[width=0.24\textwidth]{figs/adR4em1adS7em1Chi5em1_ra090_image08.png}
  \caption{\label{fig:doublet090} The formation of a doublet in a shear
  flow.  The reduced area is $\Delta A = 0.9$, the Hamaker constant is
  $\mathcal{H}=0.7$, the separation distance is $\delta = 0.4$, and the
  shear rate is $\chi=0.5$.}
\end{figure}

We characterize the effect of the Hamaker constant and shear rate by
determining a critical Hamaker constant, $\mathcal{H}_c$, which
determines if the vesicles form a doublet or separate.  We fix the
separation distance to $\delta = 0.4$ and the vesicle reduced area to
$\Delta A = 0.9$.  In the left plot of
Figure~\ref{fig:sflow_phase_diagram}, we plot the minimum distance
between the vesicles as a function of time for various Hamaker
constants.  The red curves correspond to Hamaker constants that are too
small to bind the vesicles into a doublet, and the blue curves
correspond to Hamaker constants that are sufficiently large to form a
doublet.

\begin{figure}
  \includegraphics[height=0.35\textwidth]{figs/shear_adR4em1Chi1e0_ra-090.pdf}
  \includegraphics[height=0.35\textwidth]{figs/shear_adR4em1_ra090_phaseDiagram.pdf}
  \caption{\label{fig:sflow_phase_diagram} {\em Left}: Distance between
  a pair of vesicles in a planar shear flow with shear rate $\chi=0.5$,
  and separation distance $\delta = 0.4$.  The different lines
  correspond to a linear spacing of Hamaker constants ranging from
  $\mathcal{H}=0.1$ and $\mathcal{H}=1.0$.  {\em Right}: Phase diagram
  of vesicle hydrodynamics in a shear flow.  At a fixed reduced area
  $\Delta A=0.90$ and $\delta = 0.4$, the critical Hamaker constant
  $\mathcal{H}$ for binding of two vesicles in a shear flow depends on
  the shear rate $\chi$.}
\end{figure}

Finally, we investigate the rheology of a suspension of a doublet by
computing the effective viscosity of a doublet and compare it to the
effective viscosity of a single tank-treading vesicle. The effective
viscosity is defined as the viscosity of a homogeneous Newtonian fluid
with the same energy dissipation per macroscopic element of fluid.  In a
simple shear flow, the intrinsic viscosity, $[\mu]$ is
\begin{align*}
  [\mu]:= \frac{\mu_{\mathrm{eff}} - \mu_0}{\phi \mu_0} = 
  \frac{1}{\chi \mu_0 T} \int_{T_i}^{T_e} 
  \langle \sigma_{12} \rangle dt,
\end{align*}
where
\begin{align*}
  \langle \tau \rangle = \frac{1}{|\omega|} \int_{\gamma}
    \xxi \otimes \xx ds,
\end{align*}
$\phi$ is the area fraction of vesicles, $\tau$ is the stress due to the
vesicles, and $\langle \cdot \rangle$ is the spatial average.  In the
left plot of Figure~\ref{fig:shearIntrinsicViscosity}, we compare the
intrinsic viscosity of a single tank-treading vesicle to a doublet.  To
validate our simulations, we superimpose (black marks) the intrinsic
viscosity calculated by Ghigliotti {\em et
al.}~\cite{GhigliottiBibenMisbah2010_JFM}.  The presence of the doublet
significantly increases the intrinsic viscosity at all the reduced
areas.  To further characterize the effect of adhesion, in the right
plot of Figure~\ref{fig:shearIntrinsicViscosity} we decompose the
intrinsic viscosity into the contributions from the bending and tension
(blue), and the contribution from the adhesion (red).  We also
superimpose twice the intrinsic viscosity of a single tank-treading
vesicle (dashed curve) to demonstrate that the bending and tension of
the doublet behave similarly, but not identically, to a dilute
suspensions of non-adhering tank-treading vesicles.  We see that the
effect of the adhesion on the intrinsic viscosity is largest for
vesicles with small reduced areas.

\begin{figure}[htp]
  \centering
  \includegraphics[width=0.45\textwidth]{figs/shear2Ves_adR4em1adS7em1Chi5em1.pdf}
  \includegraphics[width=0.45\textwidth]{figs/doublet_decomp.pdf}
  \caption{\label{fig:shearIntrinsicViscosity} {\em Left}: The intrinsic
  viscosity of a tank-treading vesicle (blue) and a doublet (red).  The
  shear rate is $\chi = 0.5$, the Hamaker constant is $\mathcal{H} =
  0.7$, and the separation distance is $\delta = 0.4$.  The black marks
  denote intrinsic viscosity values computed by Ghigliotti {\em et
  al.}~\cite{GhigliottiBibenMisbah2010_JFM} (cf.~Figure 5).  {\em
  Right}: The decomposition of the intrinsic viscosity into the
  contributions for the bending and tension (blue) of the vesicles, and
  the contribution from the vesicle adhesion (red).  Also included is
  twice the intrinsic viscosity of a single tank-treading vesicle
  (dashed black line).}
\end{figure}

@@

Flormann {\it et al.} studied the clustering and packing of vesicles in
free suspension~\cite{FlormannAouane2017_SciReports}.  In engineering
applications of vesicle emulsions often smaller vesicles are enclosed in
a big vesicle, as shown in Figure~\ref{fig:relaxationManyVes}, where
the left panel shows an initial condition of five vesicles suspended
inside of a larger vesicle.  The second panel from left shows the
membrane configuration and flow streamlines at an intermediate time.
There is no imposed flow, so the dynamics are governed entirely by the
adhesion, bending, and tension forces.  In the absence of confinement,
we would expect the vesicles to align themselves as a rouleaux with the
membranes forming a parachute, male-female, or S-shape
configurations~\cite{FlormannAouane2017_SciReports}.  However, because
of the bounding vesicle and the dense packing, the interior vesicles
align themselves differently.  In addition, several parts of the
bounding membrane closely follow the interior membranes in order to
minimize the adhesion energy.   As a result of force balance the
equilibrium shape of the bounding vesicle is far from a symmetric shape
while the interior smaller vesicles are of similar shape to each other.
We also plot the tension (thrid) and adhesion (fourth) of the final
equilibrium configuration.  We remark that the adhesion force is all
attractive on all the vesicle membranes while large positive membrane
tension is found only on the bounding vesicle.  In particular, the
interior vesicles form a flat contact region with a large negative
tension.  We are now conducting simulations of such vesicle
configuration under a planar shear flow.
